% English
\null\vfil
\begin{otherlanguage}{english}
\begin{center}\textsf{\textbf{\abstractname}}\end{center}

    \noindent PharMe is a service that aims to bring \glsa{pgx} and its
    potential advantages to the treatment of patients by providing them with
    personalized \gls{pgx} insights. To be useful to patients, these insights
    need to be communicated comprehensively. Researching, curating and
    administering the necessary information is a time intensive process
    conducted by \glsa{pgx} experts. This process is facilitated by the
    Annotation Interface researched and implemented in this thesis. The primary
    used method modularizes information to be made up of predefined textual
    components rather than free text. This increases consistency,
    maintainability, and curation efficiency of information, while also laying
    the foundation for support of multilingualism: By merely translating its
    components, all current and future information is available in the
    respective language. The presented methods attend PharMe's need for
    administration of information and provide insights towards establishing
    systems to solve similar problems for other health-related informative
    resources.

\end{otherlanguage}
\vfil\null

% German
\null\vfil
\begin{otherlanguage}{ngerman}
\begin{center}\textsf{\textbf{\abstractname}}\end{center}

    \noindent PharMe ist ein Service mit dem Ziel, PGx und seine potenziellen
    Vorteile in die Behandlung von Patient*innen einzubringen, indem er ihnen
    personalisierte pharmakogenomische Erkenntnisse liefert. Um für die
    Patient*innen nützlich zu sein, müssen diese Erkenntnisse verständlich
    vermittelt werden. Das Recherchieren, Zusammenstellen und Verwalten der
    erforderlichen Informationen ist ein zeitintensiver Prozess, der von
    PGx-Expert*innen durchgeführt wird. Dieser Prozess wird durch das
    Annotation Interface, das mit dieser Arbeit erforscht und implementiert
    wird, erleichtert. Die primär verwendete Methode modularisiert die
    Informationen, so dass sie aus vordefinierten Textkomponenten anstatt aus
    freiem Text bestehen. Dies erhöht die Konsistenz und Effizienz der
    Informationspflege und schafft gleichzeitig die Grundlage für die
    Unterstützung mehrerer Sprachen: Bereits die Übersetzung der Komponenten
    genügt, um alle aktuellen und zukünftigen Informationen in der jeweiligen
    Sprache zur Verfügung zu stellen. Die vorgestellten Methoden erfüllen
    PharMe's Erfordernis an die Verwaltung von Informationen und bieten
    Einblicke in die Entwicklung von Systemen zur Lösung ähnlicher Probleme für
    andere gesundheitsbezogene Informationsressourcen.

\end{otherlanguage}
\vfil\null
